Sperm should evolve to make female meiosis fair.

### 2014-05-21 writeup_and_notes/signif_statement.tex
yb got sig statement down to 149 words, still need to cut 30

Evolutionary biologists are fascinated by the oppotunities for benefits, costs, and opportunities for conflict during meiosis, however,little attention has been directed at the odd fact that fertilization by sperm is required to complete female meiosis in many species of animal. At first sight this seems strange as it seems to presents a dangerous opportunity for selfish alleles in sperm to distort the outcome of female meiosis in their own favor creating the opportunity for conflicts between male and female genomes. In this paper, through mathematical modeling, we show that in fact quite the opposite is true, sperm are usually selected to ensure the fairness of meiosis. Intuitively this comes about because selfish sperm have to immediately suffer any ill consequences of their actions in the resulting zygote, and so it is strongly in sperm's interests to keep genomes from gaming female meiosis. We hypthosize that the wide-spread requirement of fertilization to complete female meiosis may represent a long-term collaboration between male and female gametes to ensure fairness and the production of high quality gametes.

>>>>

\documentclass{article}
\usepackage{graphicx}

\begin{document}
Female meiosis is the time when alleles can directly modify their transmission. Benefits to alleles that succeeds in this competition and collateral damage to organismal fitness likely shaped basic features of female reproduction. However, the observation that fertilization is often required for the completion of female meiosis appears incompatible with this view. In fact, this fertilization requirement seems to presents an opportunity for sperm-acting alleles to distort the in their own favor, creating an additional arena for conflict. We show that, counterintuitively, the opposite is true, sperm are usually selected to ensure that female meiosis is fair. Intuitively this comes about because sperm immediately suffer the company they keep in the zygote, and it is in sperm's interests to enforce fairness in female meiosis. The wide-spread requirement of fertilization to complete female meiosis may represent a long-term collaboration between sperm and mothers to ensure the production of fit offspring.
\end{document}
